\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-03-29 paper.qmd
Introduction complete

```{=html}
<!-- 
  Notes
* Use American spelling
-->
```

@@

 - What's the problem:

@@

  Relevant high-d vis, NLDR history

@@

```{r}
#| label: NLDR-variety
#| echo: false
#| fig-cap: "Six different NLDR representations of the same data. Different techniques and different parameter choices are used. Which is the most accurate representation of the structure in high dimensions?"
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%
# 2D layouts from different NLDR techniques and different hyperparameter choices applied for the PBMC3k dataset: (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (d) UMAP (n_neighbors = 15, min_dist = 0.03), (e) UMAP (n_neighbors = 84, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn = 2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (m) TriMAP (n_inliers = 30, n_outliers = 4, n_random = 10), (n) TriMAP (n_inliers = 5, n_outliers = 2, n_random = 2), (p) PaCMAP (n_neighbors = 10, init = pca, MN_ratio = 0.5, FP_ratio = 2), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5), (r) PaCMAP (n_neighbors = 5, init = pca, MN_ratio = 0.1, FP_ratio = 1), and (s) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.1, FP_ratio = 1). Is there a best representation of the original data or are they all providing  equivalent information?
nldr1 + nldr2 + nldr3 + nldr4 +
  nldr5 + nldr6 + nldr7 + nldr8 +
  plot_layout(ncol = 4)
```

## Method

@@

## Applications

@@

## Discussion

>>>>

<!-- 
Check-list before submission
* Is it all American spelling
* Spelling checked generally
* Code all runs given fresh workspace
* Code has a readme, explaining how the paper results are reproduced
-->

@@

Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional representation of high-dimensional data. Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008], uniform manifold approximation and projection (UMAP) [@Leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@article03], large-scale dimensionality reduction Using triplets (TriMAP) [@article02], and pairwise controlled manifold approximation (PaCMAP) [@Yingfan2021]. However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety). The dilemma for the analyst is then, **which representation to use**. The choice might result in different procedures used in the downstream analysis, or different inferential conclusions. The research described here provides new visual tools to aid with this decision.

<!-- - What's the problem:

@@

  Relevant high-d vis, NLDR history
-->

@@

```{r}
#| label: fig-NLDR-variety
#| echo: false
#| fig-cap: "Six different NLDR representations of the same data. Different techniques and different parameter choices are used. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice. Would they make different decisions downstream in the analysis depending on which version seen? Which is the most accurate representation of the structure in high dimensions?"
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%
# 2D layouts from different NLDR techniques and different hyperparameter choices applied for the PBMC3k dataset: (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (d) UMAP (n_neighbors = 15, min_dist = 0.03), (e) UMAP (n_neighbors = 84, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn = 2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (m) TriMAP (n_inliers = 30, n_outliers = 4, n_random = 10), (n) TriMAP (n_inliers = 5, n_outliers = 2, n_random = 2), (p) PaCMAP (n_neighbors = 10, init = pca, MN_ratio = 0.5, FP_ratio = 2), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5), (r) PaCMAP (n_neighbors = 5, init = pca, MN_ratio = 0.1, FP_ratio = 1), and (s) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.1, FP_ratio = 1). Is there a best representation of the original data or are they all providing  equivalent information?
nldr1 + nldr2 + nldr3 + nldr4 +
  nldr5 + nldr6 + nldr7 + nldr8 +
  plot_layout(ncol = 4)
```

The paper is organised as follows. @sec-background provides a summary of the literature on NLDR, and high-dimensional data visualization methods. @sec-method contains the details of the new methodology, including simulated data examples. Two applications illustrating the use of the new methodology for bioinformatics and image classification are in @sec-applications. Limitations and future directions are provided in @sec-discussion.

## Background {#sec-background}

## Method {#sec-method}

@@

## Applications {#sec-applications}

@@

## Discussion {#sec-discussion}

### 2024-03-29 paper.tex
Introduction complete

\begin{itemize}
\tightlist
\item
  What's the problem:
\end{itemize}

Non-linear dimension reduction being used to summarise high-dimensional
data.

\begin{itemize}
\tightlist
\item
  Summary of literature
\end{itemize}

Relevant high-d vis, NLDR history

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/NLDR-variety-1.pdf}

@@

\caption{Six different NLDR representations of the same data. Different
techniques and different parameter choices are used. Which is the most
accurate representation of the structure in high dimensions?}

@@

\section{Method}\label{method}

@@

\section{Applications}\label{applications}

@@

\section{Discussion}\label{discussion}

>>>>

Non-linear dimension reduction (NLDR) is popular for making a convenient
low-dimensional representation of high-dimensional data. Recently
developed methods include t-distributed stochastic neighbor embedding
(tSNE) \citep{Laurens2008}, uniform manifold approximation and
projection (UMAP) \citep{Leland2018}, potential of heat-diffusion for
affinity-based trajectory embedding (PHATE) algorithm \citep{article03},
large-scale dimensionality reduction Using triplets (TriMAP)
\citep{article02}, and pairwise controlled manifold approximation
(PaCMAP) \citep{Yingfan2021}. However, the representation generated can
vary dramatically from method to method, and with different choices of
parameters or random seeds made using the same method
(Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is then,
\textbf{which representation to use}. The choice might result in
different procedures used in the downstream analysis, or different
inferential conclusions. The research described here provides new visual
tools to aid with this decision.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-variety-1.pdf}

@@

\caption{\label{fig-NLDR-variety}Six different NLDR representations of
the same data. Different techniques and different parameter choices are
used. Researchers may have seen any of these in their analysis of this
data, depending on their choice of method, or typical parameter choice.
Would they make different decisions downstream in the analysis depending
on which version seen? Which is the most accurate representation of the
structure in high dimensions?}

@@

The paper is organised as follows. Section~\ref{sec-background} provides
a summary of the literature on NLDR, and high-dimensional data
visualization methods. Section~\ref{sec-method} contains the details of
the new methodology, including simulated data examples. Two applications
illustrating the use of the new methodology for bioinformatics and image
classification are in Section~\ref{sec-applications}. Limitations and
future directions are provided in Section~\ref{sec-discussion}.

\section{Background}\label{sec-background}

\section{Method}\label{sec-method}

@@

\section{Applications}\label{sec-applications}

@@

\section{Discussion}\label{sec-discussion}
